%% Copyright 1998 Pepe Kubon
%%
%% `ack.tex' --- aknowledgments for thes-full.tex, thes-short-tex from
%%               the `csthesis' bundle
%%
%% You are allowed to distribute this file together with all files
%% mentioned in READ.ME.
%%
%% You are not allowed to modify its contents.
%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%       Acknowledgment
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\prefacesection{Acknowledgments}

It is a great honor and pleasure for me to have my Master study in Simon Fraser University. I would like to give a big thank to Dr. Robert Cameron for all of his earnest instructions. I learnt a lot from him during these two years.

I would also like to thank Dr. Nick Sumner for his deep knowledge on LLVM and his great patience in assisting my thesis work. I would like to thank Dr. Thomas Shermer and Dr. Fred Popowich for early feedback on my thesis draft. I would like to give another big thank to Ken Herdy, who gave me a lot of support in those late nights in the lab. Nigel Medforth helped me with the experiments on icgrep as well as many instructions on the Parabix tool chain. Linda Lin helped me on Parabix tools too.

I would like to thank all the committee members for their precious time spent on my thesis. I would like to thank Simon Fraser University and Dr. Cameron again for providing me this awesome chance of studying computing science in Canada as well as providing financial support. I would like to thank my parents and all my friends in Canada that encouraged me and supported me during this journey.












